\section{The DP4 library: what we are shipping}\label{sec:dp4}
This section inventories the components we contribute to the DP4 assay run and what each one is
for. DP4 as a whole is the larger fourth PepSeq library; the pieces assembled here are the
designed and control constructs this project adds to it. Each component targets one of the three
questions from Section~\ref{sec:intro}, and all of them are emitted in the same constant-length
construct (Section~\ref{sec:tiling}) and the same 8-column annotated format, so they can be merged into
one ordered synthesis file (Section~\ref{sec:assembly}).

\begin{table}[h]\centering\small
\begin{tabular}{llcl}
\toprule
component & what it is & constructs & status \\
\midrule
C1 known antibody, whole epitope & full-epitope scaffolds, top-$n$/epitope & depth-set & ranked \\
C2 known antibody, single island & 87 island contigs, each island alone & depth-set & ranked, per island \\
C3 scaffolds, polyclonal tiling & top-$n$/window, no antibody & depth-set & selected, ungated \\
C4 linear tiled-30mer controls & bare tiled peptides & 2034 & built \\
C5 metric-space sampling & designs spread across metrics & \num{3000} & sampled \\
C6 scaffolded-epitope controls & island$\to$Ala $+$ scaffold-disruption & $\sim$3k & built \\
\bottomrule
\end{tabular}
\caption{The DP4 components assembled here. C1 and C2 are the two halves of the islands
experiment (whole epitope vs.\ one island at a time); C6 tests the same question by mutation.
C1--C3 are selected (ranked on the composite, Section~\ref{sec:selection}); the shipped depth
(top-$n$ per group) is set at assembly from the peptide budget, so their final counts settle then.
C5 is set at \num{3000} designs; C4 is the \num{2034} tiles of Section~\ref{sec:tiling}.}
\label{tab:dp4components}
\end{table}

\subsection{C1 --- known-antibody epitopes, whole}\label{sec:dp4c1}
For each mAb epitope we ship the best few scaffolds that present the \emph{complete} epitope ---
all of its islands held together in their native geometry, the way the antibody actually sees it.
These are the designs whose binding we can score against a ground truth, since the antibody is
known; they continue the DP3 set that gave us the binding data of Section~\ref{sec:experimental}
and are the comparator for the single-island arm below. Selection ranks on the re-weighted
composite with no hard gate (Section~\ref{sec:selection}); because the antibody is known, the
accessibility term is the \emph{real} clash \texttt{af3\_n\_clash\_res} rather than the cylinder
surrogate. Whole-epitope presentation is straightforward for the \num{13} single-island epitopes
but hard for multi-island ones --- preserving the relative geometry of separated islands is
exactly where scaffolding struggles (Section~\ref{sec:intro}, the \texttt{2XQB} failure) --- which
is the motivation for shipping C2 alongside it. C1 is regenerated natively at the \num{103}-mer assay
length (Section~\ref{sec:production}) rather than truncated from Lawson's \num{104}-mers, so C5 and C6,
which derive from the C1 pool, inherit the corrected length.

\subsection{C2 --- known-antibody epitopes, one island at a time}\label{sec:dp4c2}
The single-island component scaffolds each island of the \num{46} multi-island epitopes
\emph{individually}: one island per design, presented alone. This is the direct test of Q1
(Section~\ref{sec:islands}) --- does an antibody need both islands of its epitope, or does one
suffice? Shipping C1 (whole epitope) and C2 (each island alone) for the same antibodies lets the
assay deconvolve which island, if any, carries the binding. The per-island production run
(Section~\ref{sec:perisland}) covers \num{87} scaffolded islands (\num{5} size-1 islands are
skipped as un-presentable); at top-5 per island that is on the order of \num{435} constructs
(Table~\ref{tab:dualisland}). Selection is identical to C1 (composite, no gate, real-clash
accessibility), ranked per (\texttt{id}, island) so each island competes on its own.

\subsection{C3 --- scaffolded epitopes from polyclonal tiling}\label{sec:dp4c3}
For antigens with no known antibody we tile the whole protein into overlapping windows and
scaffold each one (Section~\ref{sec:tiling}), then keep the top 5 designs per window by the
composite. This is the polyclonal arm (Q3): we do not know where the real epitope is, so we cover
the antigen densely and let the assay find it. This is the one component where the antibody-clash
filter cannot be computed, so accessibility is judged entirely by the native-aware cylinder
surrogate (Section~\ref{sec:noab}) --- the metric the DP3 binding data singled out as the
strongest predictor (Section~\ref{sec:whatpredicts}), which is why it now carries the most weight
in selecting these designs. Across the three tiled antigens (1D2K, 6M0J, 4WAT) this is
\num{951}+\num{450}+\num{780} top-5 designs (Table~\ref{tab:tilecount}).

\subsection{C4 --- linear tiled-30mer controls}\label{sec:dp4c4}
The linear controls are the bare epitope peptides the scaffolded designs are meant to beat:
unscaffolded 30-mers (model \texttt{(none)}) tiled across the antigen, presenting the sequence
with none of the conformational constraint a scaffold imposes. They are the baseline for the
central scaffold-versus-linear question (Section~\ref{sec:intro}): if conformational presentation
matters, a scaffolded design should bind its antibody more reliably than the corresponding linear
tile. We tile the gap-free PDB FASTA chain at a 30-residue window, step 6
(Section~\ref{sec:tiling}), giving \num{2034} tiles over \num{59} antigens --- the \num{56}
known-antibody targets plus the \num{3} polyclonal antigens
(\texttt{data/libraries/dp4\_tiled30mers\_fasta.csv}). The three non-standard cases excluded from the
antibody arms (\texttt{2h32} the pre-B-cell receptor, \texttt{4xwo} low-yield, \texttt{7a3t} too-small;
Section~\ref{sec:dp4c6}) are dropped here as well, so the linear controls stay consistent with the
scaffold arms they are meant to be compared against. This component is built.

\subsection{C5 --- metric-space sampling}\label{sec:dp4c5}
The DP3 binding data is a weak prior precisely because every assayed design already passed the
four filters, so the metrics barely varied (Section~\ref{sec:experimental}). C5 fixes that by
design: instead of shipping only the top designs, we deliberately sample designs \emph{spread
across} the metric space --- including ones the current filters would reject --- so that epitope
RMSD, accessibility, and PAE each take a wide range of values. Here accessibility is the cylinder
(not the real clash): the point of this component is to learn the predictors that transfer to the
no-antibody setting, so we sample along the same surrogate we will have to rely on there. Reading
binding off this spread is what lets us learn which metrics actually predict enrichment and
finally fit the composite's weights (Q2, Section~\ref{sec:open}), rather than re-deriving the
filters we started from. The sampling is implemented (\texttt{scripts/stage06\_sample\_c5.py}): a
farthest-point (max--min) spread over the four standardized scoring axes --- epitope RMSD, epitope
PAE, overall RMSD, and the native-aware cylinder --- sampled per mAb over the \num{56} assayable
epitopes. We ship \num{3000} designs (about \num{54} per mAb); the sample spans \SIrange{89}{99}{\percent}
of each axis's full range, i.e.\ deliberately including the low-quality tail the filters reject.

\subsection{C6 --- scaffolded-epitope controls}\label{sec:dp4c6}
For the best \num{20} designs of each of the \num{56} mAbs we alanine-scan the epitope islands:
every residue of island~1 mutated to alanine, and (where the epitope has a second island) every
residue of island~2. The unmutated designs are not part of this component --- they are already
shipped as C1 --- so C6 is the controlled comparison against them. Mutating an island to alanine
strips its side-chain chemistry while leaving the fold intact, so if binding drops when island~1 is
alanine-substituted the antibody requires island~1, and likewise for island~2. This is the
mutation-based read of the islands question (Q1), complementary to C2's presentation-based read
(each island scaffolded alone): C2 asks whether one island \emph{suffices}, C6 asks whether each
island is \emph{necessary}. The mutations are string substitutions on the existing designs'
case-encoded sequence (epitope residues uppercase, scaffold lowercase, each contiguous uppercase run
an island), implemented in \texttt{episcaf\_pipeline/scaffolded\_epitope\_controls/} (a port of the
DP3 scheme). Over the best designs of each of the \num{56} mAbs, with one island-mutant for
single-island epitopes and two for multi-island ones, plus the scaffold-disruption variant below,
this is on the order of \num{3000} constructs at the current selection depth.

\paragraph{Scaffold-disruption control.} A companion negative control mutates the \emph{scaffold}
(not the epitope) in each design: if binding is epitope-driven, perturbing the scaffold should leave
it intact, so a design that survives this but collapses under an island alanine-substitution is
behaving as intended. We insert a fixed structure-disrupting hexamer (\texttt{PPDDGG}) into four
lowercase (scaffold) windows, each required to lie at least four residues from any epitope position,
so the epitope is untouched and the perturbation is confined to the scaffold; placement is seeded for
reproducibility. We ship this four-window variant (\texttt{scaffoldMutX4}) --- the stronger
perturbation carried over from DP3 --- and drop the single-window one to save space. Because island
membership and the epitope-avoidance constraint are both read directly from the case-encoding, no
separate residue-index map is needed. When a scaffold cannot fit four disjoint windows, we
\emph{degrade the count} (\texttt{scaffoldMutX3}, \texttt{X2}, \texttt{X1}) rather than drop the control,
so a design loses its scaffold control only if it cannot place even a single hexamer $\ge\!4$ residues
from the epitope. In the current build the four-window arm covers \num{1034} designs, with \num{53} at
three windows and \num{33} at two, so \emph{every} one of the \num{1120} bases carries a scaffold
control; the island-alanine arms likewise cover all bases.

\medskip
\noindent The controls follow the DP3 mutation scheme (reference and port in
\texttt{episcaf\_pipeline/scaffolded\_epitope\_controls/}); the DP4 changes are all-residue island
alanine (not every other) and the four-window scaffold variant only. The base set is the same top-$n$
per mAb used for C1.

\paragraph{The known-antibody set (\num{56}).} The antibody-based components (C1, C2, C5, C6) share one
exclusion set applied consistently: from the \num{59} DP3 mAb epitopes we drop three, leaving \num{56}.
Two are the assay drops of Section~\ref{sec:experimental} (\texttt{4xwo}, low antibody yield;
\texttt{7a3t}, a \num{4}-residue epitope too small to present). The third, \texttt{2h32}, is excluded
because it is not a conventional antibody:antigen complex at all --- it is the pre-B-cell receptor, which
engages its ligand differently, so it is not a valid test case for antibody-epitope scaffolding; it
entered the DP3 set through the inclusion criteria inadvertently. The polyclonal tiling (C3) is over
different antigens and is unaffected.

\subsection{8VDL --- a PfEMP1 conserved-epitope arm}\label{sec:dp4_8vdl}
One further arm scaffolds a conserved epitope from a different antigen entirely: the EPCR-binding
surface of the \emph{Plasmodium falciparum} PfEMP1 CIDR$\alpha$1.4 domain (crystal structure
\texttt{8VDL}; Reyes et al., \emph{Nature} 636:182--189, 2024). PfEMP1 is the hypervariable parasite
surface antigen that drives severe malaria and escapes immunity by antigenic variation, but the
residues its CIDR domain uses to grip the host receptor EPCR cannot vary freely --- so a construct that
presents that conserved contact surface is a candidate for eliciting \emph{broadly} reactive,
variant-transcending antibodies. The paper identifies the conserved, EPCR-binding-critical residues on
which its broadly reactive antibodies depend: the double-phenylalanine motif \texttt{F655}/\texttt{F656}
and a supporting glutamate \texttt{E666}. We scaffold the contiguous chain~C window
\numrange{651}{670} (\texttt{FDSFFFQVIYKFNEGEAKWN}) that spans those hotspots, and the hotspots alone,
as the two epitope definitions below. Because the crystal also contains the cognate C7 antibody
(chains H/L), this is a known-antibody target and the real clash term applies, exactly as for C1/C2.

We scaffold it two ways and keep the best \num{10} of each: \texttt{epitope20} fixes the whole
\num{20}-residue island (the strongest constraint, the direct analog of C1), while \texttt{hotspots}
fixes only the three functional residues \texttt{F655}/\texttt{F656}/\texttt{E666} and lets the design
build everything else around them (a minimal ``hotspot graft''). Both are scaffolded into the same
\num{103}-mer construct through the same RFdiffusion3~$\rightarrow$~ProteinMPNN~$\rightarrow$~AlphaFold3
pipeline and ranked on the same composite; the fixed atoms carry the residues' native crystal
coordinates, so the hotspot arm preserves their three-dimensional arrangement however the construct
spaces them. Comparing the two in the assay asks whether the minimal hotspot cluster suffices or the
whole epitope is needed. This arm is produced as a self-contained component (\texttt{dp4\_8vdl/}) and
its top designs are merged into the library alongside the others at assembly.

\subsection{Assembly into one synthesis file}\label{sec:assembly}
All six components share the constant 103-mer construct and the eight-column annotated format
(Section~\ref{sec:tiling}), so the final step concatenates them into a single ordered library file
with global \texttt{library\_member} numbering --- a named, documented pipeline stage
(\texttt{06\_library}) so the shipped file is regenerable from its components. Selection is done
(C1--C3 ranked, C5 sampled at \num{3000}); assembly runs once the shipping depth is chosen and the
per-design sequences are recovered, as follows.

\paragraph{Recovering the design sequences (case encoding).} Assembly and the C6 controls both need,
for each selected design, its full amino-acid sequence with the epitope residues marked --- stored as
a \emph{case-encoded} string (epitope uppercase, scaffold lowercase, so each contiguous uppercase run
is one island). This has to be rebuilt: our selection tables identify designs by a hash token, but the
epitope-position annotation lives in the design table \texttt{dp2} under a different key that was not
carried forward, so the token$\to$design mapping is missing for all but the assayed subset. We recover
it through the one key that always survives --- the design's own sequence: for each selected design we
read its chain-A sequence from its structure, match it to its \texttt{dp2} row, and uppercase the
epitope chunk-span positions (\texttt{scripts/case\_encode\_selected.py}, run once over the selected
set on the cluster). This yields the case-encoded sequences both C6 and assembly consume, and writes
the mapping back out so it need not be rebuilt again.

The assembled peptide library is then encoded to DNA oligos for synthesis with the LadnerLab
\texttt{oligo\_encoding} tools (\url{https://github.com/LadnerLab/Library-Design}), using the same
updated codon-weight table as DP3
(\texttt{/tgen\_labs/Immunology/ekelley/DP3\_PepSeq\_Library\_Design/new\_codon\_weights/}). The
encoding is the final synthesis-prep step; it is external tooling, so the pipeline ends at the
annotated peptide file and hands off to it.
