\section{Motivation: scaffolding B-cell epitopes}\label{sec:intro}

\subsection{Why scaffold epitopes}
This project is about \emph{B-cell epitopes}: the regions of an antigen an antibody recognizes. An
antibody reads those regions by their three-dimensional shape, and many of the most interesting ones
are discontinuous in sequence but contiguous in space. The contacted residues are scattered along the
antigen sequence, yet they fold together into a single surface patch. A linear representation cannot
hold that.

High-throughput serology platforms such as PepSeq tile an antigen into short, overlapping linear
peptides. This works well for continuous, sequence-local epitopes. It cannot work for a discontinuous
one: a linear peptide reconstitutes neither a surface assembled from sequence-distant residues nor the
native conformation the antibody evolved against.

The central question of this project is whether \emph{scaffolding} an epitope produces a stronger,
cleaner binding signal in the PepSeq assay than linear tiling does. Embed the epitope residues in a
folded protein and you hold them in the conformation the antibody actually sees, and you can present
distinct epitopes as separable, individually testable designs. If conformational presentation matters,
a scaffolded design should bind its cognate antibody more reliably than the corresponding linear
peptide. That is the test. The longer-term goal is to carry PepSeq-style assays past linear peptides to
conformational, discontinuous epitopes.

\subsection{The approach}
We extract the epitope --- the residues an antibody contacts --- from an antibody:antigen crystal
structure, embed those residues into a \emph{de novo} scaffold with RFdiffusion, assign a sequence with
ProteinMPNN, and ask AlphaFold3 whether the design folds with the epitope held in its native geometry.
Designs are kept or rejected on AlphaFold3 confidence and geometry-preservation metrics
(Section~\ref{sec:methods}). The production pipeline --- the Nextflow orchestration, the RFdiffusion
runs, the AlphaFold3 evaluation --- originates in Lawson Woods' \texttt{episcaf-experiments} repository.
Reproducing it, then extending it, is where this work starts.

\subsection{Patches and islands}
A single antibody:antigen complex can yield more than one epitope \emph{patch} (\texttt{0P},
\texttt{1P}, and so on), each a distinct cluster of antibody-contacting residues from the same
structure. A patch is in turn made of one or more \emph{islands}: the spatially contiguous runs of
epitope residues in three-dimensional space. The distinction matters because it predicts difficulty. A
single-island patch can be enforced locally and tends to scaffold well. A multi-island patch has to
preserve the relative position and orientation of spatially separated regions, which is much harder.

One case makes the point. PDB \texttt{2XQB}, patch \texttt{0P}, is a 56-residue epitope split across two
separated islands, and across \num{2624} RFdiffusion trials zero designs passed the downstream filters.
A $0\%$ rate over thousands of attempts is a real computational failure, not undersampling. The cause is
procedural: contigs (the constraints RFdiffusion is given) fix epitope residue identity and ordering,
but they do not strongly constrain the distances and orientations \emph{between} islands, so the model
can satisfy the contig and still lose the geometry an antibody needs.

\subsection{Design pools and the three DP4 questions}
The project spans two PepSeq iterations. DP3 is the previous round: the known-antibody mAb set, where a
solved complex gives a ground truth for ``good.'' DP4 is the library we are assembling now --- several
thousand 103-mers that \emph{mostly} re-target those same mAb epitopes, and for a minority of antigens
extend to whole proteins with no known antibody. The no-antibody antigens are one component of DP4, not
the whole of it; but they are the component that forces the accessibility problem, because the
antibody-clash filter cannot be computed without an antibody. So scoring runs in two regimes:
\begin{itemize}
  \item \textbf{Known antibody} (all of DP3, most of DP4). The solved complex gives the exact epitope
  and a ground truth for ``good'': the design must hold the epitope in place and must not put scaffold
  mass where the antibody has to bind. This is where we have already run designs through the assay and
  seen success, and where scaffolded-versus-linear is most directly testable.
  \item \textbf{No antibody} (one component of DP4). Without a solved complex we do not know where the
  epitope is, so we tile the sequence into overlapping 12-residue windows (step 2) and scaffold each,
  covering the antigen densely enough to catch whatever the real epitopes turn out to be. This is the
  polyclonal arm, and the one place the native-aware cylinder surrogate (Section~\ref{sec:noab}) is
  needed --- there is no antibody to clash against.
\end{itemize}

\noindent
DP4 is built to answer three questions, and they organize the rest of this log.
\emph{(Q1) What is a minimal epitope?} How many islands and residues does binding need --- one island or
both? The island analysis (Section~\ref{sec:islands}), the whole-epitope set (Section~\ref{sec:c1set}),
and the per-island production run that tests it (Section~\ref{sec:perisland}) are aimed here.
\emph{(Q2) What predicts a successful design?} Of epitope RMSD, scaffold RMSD, PAE, and accessibility,
which actually track binding? This is what the decomposed metrics and composite scorer
(Section~\ref{sec:scoring}) are for. The first experimental answer is already in: the DP3 binding data
(Section~\ref{sec:experimental}) singles out accessibility and epitope RMSD. DP4 is built to sharpen it
by assaying designs spread across the feature space, so the scorer's weights can be fit rather than
hand-set (Section~\ref{sec:open}).
\emph{(Q3) Can it extend to the polyclonal setting?} Tiling a whole antigen with no known antibody, where
the cylinder surrogate stands in for the antibody-clash filter. These design ideas originate with the
project's earlier work; the scoring and metrics here are a refined evolution of them.

\subsection{How the rest of this log reads}
The log runs in two halves, and each section motivates the next.

\emph{Part I --- method and ground truth.} We reproduce Lawson's workflow and confirm our metrics match
his (Section~\ref{sec:methods}), tile antigens into assay constructs (Section~\ref{sec:tiling}), and
compare the RFdiffusion1 and RFdiffusion3 design sets head to head (Section~\ref{sec:rfd}). We then
rebuild scoring: the inherited filters bias selection toward rigid folds, so we decompose the metrics
into a composite that picks the best designs per epitope (Section~\ref{sec:scoring}), and recover the
antibody-accessibility filter for targets with no antibody through the native-aware cylinder surrogate
(Section~\ref{sec:noab}). The first experimental binding data closes the loop
(Section~\ref{sec:experimental}): it shows which metrics actually predict binding, and it is what sets
the scorer's weights.

\emph{Part II --- designing DP4.} With the method grounded, we turn to the next library. We ask whether
an antibody needs one epitope island or both (Section~\ref{sec:islands}), build the whole-epitope set at
the native 103-mer length (Section~\ref{sec:c1set}) and the per-island production run that lets the assay
answer the islands question (Section~\ref{sec:perisland}), and lay out the full set of DP4 components we
are shipping (Section~\ref{sec:dp4}). Open questions and next steps close the log
(Section~\ref{sec:open}).
