\section{The pipeline, and reproducing Lawson's workflow}\label{sec:methods}

\subsection{Pipeline}
RFdiffusion3 generates scaffold backbones around the epitope, ProteinMPNN re-sequences
them, and AlphaFold3 predicts the structure single-sequence
(\texttt{--norun\_data\_pipeline}, one diffusion sample, seed~1). Per-design metrics are
computed from the AlphaFold3 prediction against the designed structure. Generation code
lives in \texttt{episcaf\_pipeline/}, and the metrics, scoring, and visualization code in
\texttt{episcaf\_analysis/}.

\subsection{Reproducing Lawson's baseline}\label{sec:repro}
Before trusting any new number we re-implement Lawson's metrics and check them against his
ground truth. \texttt{compute\_metrics.py --validate} reads \texttt{dp2.parquet} (Lawson's
stored per-design values), recomputes \texttt{overall\_rmsd} and
\texttt{epitope\_chunk\_rmsd} from his ProteinMPNN PDBs and AlphaFold3 CIFs, and diffs them
against the stored values. Agreement on this diff is the precondition for everything
downstream: it confirms that our implementation of the filters matches the workflow we are
extending, so that later differences reflect the design protocol rather than a scoring bug.

\subsection{Four-filter ground truth (DP3 set)}
A design passes when all four conditions hold: overall backbone RMSD $\le \SI{2}{\angstrom}$,
epitope-chunk RMSD $\le \SI{1}{\angstrom}$, mean predicted aligned error $< 5$, and zero
antibody residues within \SI{4}{\angstrom} of non-epitope scaffold residues
(\texttt{af3\_n\_clash\_res}). Clash detection Kabsch-aligns the AlphaFold3 epitope onto the
true epitope, then checks non-epitope scaffold heavy atoms against antibody atoms. The
antibody coordinates always come from the AbDb crystal structure, not from AlphaFold3.

We summarize a design protocol by its empirical pass rate. If a protocol produces $N$
designs $s_i$ and $f(s_i)=1$ when design $i$ clears all four filters (and $0$ otherwise),
\begin{equation}\label{eq:passrate}
\hat P_{\text{pass}} \;=\; \frac{1}{N}\sum_{i=1}^{N}\mathbf{1}\!\left[f(s_i)=1\right].
\end{equation}
This is just the fraction of designs that pass, and it is the quantity we compare between
protocols and report per epitope.

\subsection{Decomposed metrics}\label{sec:decomp}
The two continuous filters, overall RMSD and mean PAE, treat the epitope and the scaffold
as a single unit. We split each into epitope- and scaffold-specific terms (epitope RMSD,
scaffold RMSD, epitope PAE, scaffold PAE), retaining the zero-clash hard filter throughout.
Decomposition lets selection reward a well-placed epitope on a flexible scaffold (or the
reverse) instead of demanding global rigidity, and it lets us ask whether the optimal
rigidity profile is epitope-dependent rather than universal (Section~\ref{sec:filtering}).
Selection over the four metrics is organized as a $2\times2$ of strict and relaxed
thresholds on (epitope, scaffold), with a flexible/flexible arm as a negative control.

\subsection{Native-antigen-aware cylinder surrogate (12-mer set)}
With no antibody available, we approximate the volume an antibody would occupy with a
cylinder fitted to the epitope. The epitope C$\alpha$ coordinates $\{x_j\}_{j=1}^{m}$ define
a plane: the centroid is $c=\tfrac{1}{m}\sum_j x_j$, and the plane normal $\hat n$ is the
smallest-variance singular vector of the centered coordinates $\{x_j-c\}$ (the antibody
approach direction), oriented to point away from the antigen body. The cylinder base is
$b = c + d_0\,\hat n$ with offset $d_0=\SI{-4}{\angstrom}$. A scaffold C$\alpha$ at position
$p$ lies inside the cylinder when, writing $v=p-b$, its axial and radial coordinates satisfy
\begin{equation}\label{eq:cyl}
0 \;\le\; v\cdot\hat n \;\le\; H
\qquad\text{and}\qquad
\bigl\lVert\, v-(v\cdot\hat n)\,\hat n \,\bigr\rVert \;\le\; R,
\end{equation}
with height $H=\SI{40}{\angstrom}$ and radius $R=\SI{16}{\angstrom}$. The plain count is the
number of non-epitope scaffold C$\alpha$ satisfying~\eqref{eq:cyl}; the native-aware count
additionally drops any such atom within \SI{1}{\angstrom} of a native-antigen heavy atom,
i.e.\ sitting where the native antigen already sits, which a real antibody tolerates. The
geometry core is \texttt{episcaf\_analysis/native\_cylinder\_core.py} (self-tested).

\subsection{Composite scorer}
One config-driven scorer (\texttt{episcaf\_analysis/score.py}) ranks designs in four steps:
(i)~\textbf{gate} on epitope-chunk RMSD; (ii)~transform each metric to a higher-is-better
score within a scope (pooled, or per antigen); (iii)~\textbf{weighted sum}; and (iv)~keep
the \textbf{top-$k$ per epitope}. Concretely, each design $d$ receives a score
\begin{equation}\label{eq:score}
S(d) \;=\; \sum_{m} w_m\,\varphi_m\!\bigl(x_m(d)\bigr),
\end{equation}
a weighted sum over metrics $m$ of a per-metric activation $\varphi_m$ that maps the raw
value $x_m$ to a higher-is-better score. With $\varphi$ a percentile the score is rank-based
and depends on the cohort; with a sigmoid,
\begin{equation}\label{eq:sigmoid}
\varphi(x) \;=\; \frac{1}{1+\exp\!\bigl(s\,k\,(x-x_0)\bigr)},
\qquad s=+1\ \text{if lower is better (else } -1),
\end{equation}
it is an absolute, saturating score with a tunable midpoint $x_0$ and steepness $k$. In this
form the scorer is literally a single-layer perceptron with hand-set weights (feed-forward,
no backpropagation): the activations are the per-metric nonlinearities, the $w_m$ are the
weights, and ``fitting'' it (Section~\ref{sec:open}) means learning the $w_m$ (and the
$x_0,k$) from data. The dials live in \texttt{episcaf\_analysis/presets.py}:

\begin{center}
\begin{tabular}{lc}
\toprule
metric & weight \\
\midrule
epitope-chunk RMSD & 0.35 \\
epitope PAE        & 0.25 \\
accessibility      & 0.25 \\
overall RMSD       & 0.15 \\
\bottomrule
\end{tabular}
\end{center}

\noindent
The weights sum to one (renormalized over whichever metrics a given table actually
contains). The only term that changes between the two regimes is accessibility: it is
\texttt{cylinder\_native\_aware} for the no-antibody 12-mer set and
\texttt{cylinder\_ca\_clashes} for the DP3 set, using the same cylinder geometry, with the
native-aware carve applied only where there is no real antibody to clash against. The 12-mer
preset transforms metrics per antigen and keeps the top 5 designs per epitope, while the DP3
preset transforms pooled and keeps the top 15. The weights are a hand-set starting point;
Section~\ref{sec:open} describes fitting them to the DP3 \texttt{is\_pass} label.

\subsection{Per-island scaffolding of dual-island epitopes}\label{sec:perisland}
A two-island epitope binds its antibody through two spatially separate patches, and the
native-antigen geometry alone cannot say how the binding energy is divided between them.
Scaffolding each island \emph{on its own} turns that question into an experiment: presenting
island~A in isolation, island~B in isolation, and the two together (the native arrangement)
lets the PepSeq assay deconvolve the contributions --- whether the antibody binds through
island~A, through island~B, or only through both, and whether different two-island epitopes
distribute their binding differently across their islands. This complements the
island-\emph{count} comparison of Section~\ref{sec:islands}, which asks whether two islands
are harder to scaffold than one; here we instead build the reagents to dissect a single
two-island epitope.

We define an \emph{island} as one fixed-coordinate \texttt{A}-segment of the RFdiffusion3
contig string. For example the dp2 contig \texttt{15-15/A1-16/57-57/A81-81/15-15} has two
islands, \texttt{A1-16} (16 residues) and \texttt{A81-81} (1 residue), separated by generated
scaffold of the given lengths. A \emph{dual-island epitope} is a ledger id with
\texttt{epitope\_chunks}~$=2$, i.e.\ exactly two such segments; the DP3 set contains
\num{46} of them. We take \emph{island size} to be the \texttt{A}-segment span in residues
(not the count of antibody-contact residues within it).

Following John's plan, each dual-island epitope contributes its two islands as separate
design targets, and we keep the top \num{5} designs per island under the composite scorer.
The island lengths vary widely --- from a single residue to \num{51} --- and Table~\ref{tab:dualisland}
lists both islands of all \num{46} epitopes so the run is auditable epitope by epitope. One
case needs a rule: an island of a single residue cannot be presented on its own. Rather than
drop the whole epitope, we simply \emph{skip the size-\num{1} island and still scaffold its
partner}. Five islands are skipped on this ground, one each in \texttt{2h32\_0P},
\texttt{3q1s\_0P}, \texttt{6o9i\_0P}, \texttt{7a3t\_0P}, and \texttt{7tzh\_0P} (marked in
Table~\ref{tab:dualisland}); every epitope still contributes at least one island. Of the
\num{92} islands across the \num{46} epitopes, \num{87} are scaffolded, giving \num{87}
individual-island targets and, at five designs each, \num{435} final designs. The per-island
target list (both spans, sizes, and the scaffold/skip flag) and Table~\ref{tab:dualisland}
itself are generated by \texttt{episcaf\_analysis/dual\_island\_targets.py} into
\texttt{results/dual\_island\_targets.csv} and \texttt{manuscript/tables/}, the single
reproducible source for these counts.

Each island is scaffolded into a fixed \num{103}-residue construct --- the maximum peptide
length the PepSeq assay synthesizes. We call the stretches of generated scaffold lying
N-terminal and C-terminal to a fixed island its \emph{flanks}; a single-island contig is
therefore $\langle\text{N-flank}\rangle/\,\texttt{A}a\text{-}b\,/\langle\text{C-flank}\rangle$,
and the scaffold budget the flanks share is $103-\ell$ residues for an island of length
$\ell$. We note one unexplained convention difference: Lawson's original dual-island contigs
were all \num{104} residues, not \num{103}. The change to \num{103} appears to have been made
at some point without a recorded reason; we adopt \num{103} as the assay ceiling and expect a
single terminal residue, averaged over sequence space, to be immaterial, but flag it here
because its origin is unknown.

\paragraph{Contig sampling and run size.}
A \emph{contig} pairs the fixed island(s) with one concrete scaffold-length layout, and each
contig is sampled at \num{8} RFdiffusion3 backbones $\times$ \num{8} ProteinMPNN sequences
$=\num{64}$ designs (the per-contig depth throughout this project). The run size is thus set
by how many contigs we generate. We record how Lawson reached his \num{151232}-design DP3 set,
so the scale here is never mysterious: for each epitope he held the two islands fixed and swept
the scaffold layout, keeping the two flanks balanced ($|\text{N}-\text{C}|\le 1$) and stepping
the flank length over every integer from \num{1} to about \num{20}, which makes the
\emph{inter-island gap} sweep from $\sim\!\num{46}$ down to a floor near \num{8} (the islands
are never allowed to touch). This is a deterministic enumeration, not random sampling: it
yields a median of $\sim\!\num{41}$ contigs per epitope, so $\num{59}\,\text{epitopes}\times
41 \approx \num{2363}$ contigs and $\num{2363}\times\num{64}=\num{151232}$ designs. The
quantity his contigs actually vary is the spacing between the two islands.

Our single-island targets have no inter-island gap, so that lever does not exist; the only
scaffold freedom is the \emph{flank asymmetry} --- where along the \num{103}-mer the island
sits, i.e.\ the split of the $103-\ell$ budget into N- and C-flank. We therefore generate $V$
contigs per island by \emph{randomly} sampling $V$ distinct N-flank lengths over
$[1,\,103-\ell-1]$ (under a fixed seed, so the ledger is reproducible). This is an explicitly
different diversity axis from Lawson's --- scaffold position and topology, not inter-island
geometry --- which we name here rather than leave implicit. With $V$ variants the run is
\begin{equation}\label{eq:runsize}
\underbrace{\num{87}}_{\text{islands}}\times\,V\,\times\,\underbrace{\num{8}}_{\text{RFD3}}\times\,\underbrace{\num{8}}_{\text{MPNN}}
\;=\;\num{5568}\,V \ \text{designs}.
\end{equation}
We use $V=\num{20}$, giving $\num{1740}$ contigs and $\num{111360}$ designs --- the order of
the previous DP3 set ($V\approx\num{27}$ would reach its full $\sim\!\num{150000}$). $V$ is the
\texttt{--variants} dial of \texttt{episcaf\_pipeline/build\_dual\_island\_designs.py}. The
RFD3 step is then \num{1740} contigs $\times\,\num{8}$ seeds $=\num{13920}$ array tasks,
submitted in chunks under the cluster's \texttt{MaxArraySize}.

\input{tables/dual_island_targets}
