\section{One metrics table for every decision}\label{sec:metrics}
With the per-island run designed (Section~\ref{sec:perisland}) and pushed through
RFdiffusion3 $\rightarrow$ ProteinMPNN $\rightarrow$ AlphaFold3, we extract every
per-design metric \emph{once} into a single table and make all downstream choices
(composite scoring, top-5 per island) from it, so the analysis is never recomputed. This is
our analog of Lawson's \texttt{dp2.parquet}: one row per design, written by
\texttt{scripts/stage05\_extract\_metrics.py} from the \texttt{04\_af3/outputs} predictions,
the per-island ledger (\texttt{results/dual\_island\_designs.csv}), and the native AbDb
complex PDBs.

\paragraph{Why a dedicated extractor.}
The repository already held two metric extractors, and neither fit this run directly.
\texttt{compute\_metrics.py} is the Lawson-validated antibody extractor, but its run mode is
keyed on \texttt{dp2}'s hashed tokens and the no-MPNN (RFD3$\rightarrow$AF3) layout.
\texttt{build\_12mer\_metrics.py} emits exactly the columns the composite scorer reads
(\texttt{epitope\_pae}, the cylinder terms) but parses the 12-mer naming and assumes no
antibody. Stage~05 reuses the \emph{validated primitives} of the first (backbone RMSD,
PAE, AlphaFold3 file discovery, the clash geometry) and mirrors the PAE decomposition and
cylinder calls of the second, but drives them from our per-island ledger.

The per-island set sits in a useful middle: these are the 46 dual-island \emph{mAb} epitopes,
so a solved antibody exists and the real DP3 clash filter \emph{is} available. We therefore
compute both accessibility signals --- the true antibody clash
(\texttt{af3\_n\_clash\_res}, against native chains B/C) \emph{and} the cylinder surrogate ---
which makes this run a second check of the surrogate against ground truth, on top of the DP3
check in Section~\ref{sec:noab}.

\paragraph{What each design row carries.}
The table groups into identity, geometry, confidence, and accessibility
(Table~\ref{tab:metricschema}). Geometry and confidence follow the four-filter definitions of
Section~\ref{sec:methods}; the accessibility block carries the real clash and the cylinder
side by side.

\begin{center}
\begin{longtable}{lll}
\caption{Schema of the per-design metrics table (\texttt{metrics\_dual\_island.parquet}).
One row per AlphaFold3 design.}\label{tab:metricschema}\\
\toprule
column & meaning & role \\
\midrule
\endfirsthead
\toprule column & meaning & role \\ \midrule \endhead
\multicolumn{3}{l}{\emph{Identity (which design this is)}}\\
\texttt{id}            & epitope id (e.g.\ \texttt{2h32\_0P}) & select / scope \\
\texttt{island\_index} & which of the two islands (0/1)        & select group \\
\texttt{island\_size}  & island span (residues)                & strata \\
\texttt{contig\_id}    & contig within the epitope             & provenance \\
\texttt{variant}, \texttt{n\_flank}, \texttt{c\_flank} & flank-split placement & provenance \\
\texttt{seed}, \texttt{rep}, \texttt{dldesign} & RFD3 / MPNN replica indices & provenance \\
\texttt{predID}, \texttt{af3\_dir} & design stem, output path     & join key \\
\midrule
\multicolumn{3}{l}{\emph{Geometry (backbone RMSD, Kabsch-superposed)}}\\
\texttt{epitope\_chunk\_rmsd} & design vs AF3, epitope only ($\le 1$) & gate + 0.35 \\
\texttt{overall\_rmsd}        & design vs AF3, full scaffold ($\le 2$) & 0.15 \\
\midrule
\multicolumn{3}{l}{\emph{Confidence (AlphaFold3)}}\\
\texttt{epitope\_pae}  & mean PAE over epitope residues   & 0.25 \\
\texttt{scaffold\_pae} & mean PAE over the scaffold        & diagnostic \\
\texttt{mean\_pae}     & global mean PAE ($<5$, four-filter) & diagnostic \\
\texttt{ptm}, \texttt{ranking\_score} & AF3 summary scalars & diagnostic \\
\midrule
\multicolumn{3}{l}{\emph{Accessibility (this run has both)}}\\
\texttt{af3\_n\_clash\_res} & real antibody clash count ($=0$) & ground truth \\
\texttt{cylinder\_ca\_clashes}   & scaffold C$\alpha$ in the cylinder & 0.25 (plain) \\
\texttt{cylinder\_native\_aware} & same, native volume carved out     & surrogate \\
\texttt{native\_in\_cylinder}    & native non-epitope C$\alpha$ in cylinder & test \\
\bottomrule
\end{longtable}
\end{center}

\paragraph{Definitional choices (held to the DP3-validated convention).}
Three choices fix the numbers, and we keep each at the value that makes this set comparable to
DP3. (i) Both RMSDs are computed on the \emph{backbone} (N, C$\alpha$, C, O) and
Kabsch-superposed, as in \texttt{compute\_metrics.py}, not the C$\alpha$-only variant of the
12-mer extractor. (ii) The cylinder uses \texttt{exclude\_dist}\,$=\,\SI{1}{\angstrom}$, the
value adopted for the validated 12-mer preset (Section~\ref{sec:noab}). (iii) The epitope is
the \emph{whole island span} $[\,\texttt{n\_flank},\ \texttt{n\_flank}+\texttt{island\_size}\,)$
--- the residues stage~03 actually held \textsc{fixed} during ProteinMPNN --- not the
contacts-only subset stored in the ledger. The native-frame and design-frame epitope residues
pair one-to-one because each island A-segment is contiguous in both, which we verified holds
for all \num{1740} ledger contigs ($b-a+1=\texttt{island\_size}$).

\paragraph{How it is produced.}
The extractor caches each native complex once per epitope (not once per design) and shards the
$\sim$46 epitopes across worker processes:
\begin{verbatim}
python3 scripts/stage05_extract_metrics.py \
    --af3_out_dir  runs/dual_island_rfd3/04_af3/outputs \
    --mpnn_pdb_dir runs/dual_island_rfd3/03_mpnn/mpnn_pdbs \
    --ledger       results/dual_island_designs.csv \
    --native_dir   <AbDb cleaned complex dir> \
    --out          runs/dual_island_rfd3/05_analysis/metrics_dual_island.parquet \
    --workers 16
\end{verbatim}
It writes the parquet and a CSV alongside. A companion report,
\texttt{scripts/stage05\_summarize.py}, prints the breakdowns below and writes the per-island
summary \texttt{results/dual\_island\_gate\_summary.csv} (so these numbers are regenerable,
not pasted).

\subsection{What the run yielded}\label{sec:metricsresults}
The full run scored \num{111322} designs across all 46 epitopes and 87 islands, every one with
a status of \texttt{ok}. The designs are good where it matters: median epitope-chunk RMSD is
\SI{1.20}{\angstrom} (best \SI{0.03}{\angstrom}) and median epitope PAE is $3.1$, while the
scaffold is predicted less confidently (scaffold PAE median $7.9$) --- the same
epitope-sharp / scaffold-loose split the decomposed metrics were built to exploit
(Section~\ref{sec:filtering}).

\paragraph{The design funnel.}
Table~\ref{tab:funnel} walks a design through each filter. Two thirds clear the composite
scorer's gate, and nearly half hold the epitope to within \SI{1}{\angstrom}. The strict DP3
four-filter, however, passes only $0.37\%$ --- driven down by the global PAE ($<5$) and the
zero-clash requirements, exactly the regime seen on DP3 ($<1\%$, Section~\ref{sec:islands}).
This is why per-island \emph{selection ranks by the composite score within the gate}, rather
than taking only four-filter passes, which would leave most islands with nothing to advance.

\begin{center}
\begin{tabular}{lrr}
\toprule
filter (status\,=\,ok, $n=\num{111322}$) & designs passing & \% \\
\midrule
composite scorer gate (epitope RMSD $\le \SI{2.5}{\angstrom}$) & \num{74061} & 66.5 \\
\quad epitope-chunk RMSD $\le \SI{1}{\angstrom}$              & \num{51653} & 46.4 \\
\quad overall RMSD $\le \SI{2}{\angstrom}$                    & \num{45531} & 40.9 \\
\quad global mean PAE $< 5$                                   & \num{25849} & 23.2 \\
\quad antibody clash $= 0$ (of \num{107482} with clash)      & \num{2510}  & 2.3 \\
\midrule
DP3 four-filter, all four (real clash) & \num{407} & 0.37 \\
\bottomrule
\end{tabular}
\captionof{table}{Per-design pass counts. Indented rows are the individual four-filter
thresholds (not nested); the last row requires all four at once. Regenerated by
\texttt{scripts/stage05\_summarize.py}.}\label{tab:funnel}
\end{center}

\paragraph{Top-5 per island is reachable almost everywhere.}
Ranking within the gate, \textbf{83 of the 87 islands have at least five designs to advance}.
The four that fall short (Table~\ref{tab:shortislands}) are \emph{all large islands, 21--28
residues} --- the island-size difficulty of Section~\ref{sec:islands} surfacing directly in
the deliverable: the biggest islands are the hardest to scaffold cleanly, and three of them
produce no gated design at all. For these we advance whatever clears the gate (fewer than
five, or none) and say so, rather than padding the set with designs that miss the bar.

\begin{center}
\begin{tabular}{llrr}
\toprule
epitope & island & island size (aa) & gated designs (of \num{1280}) \\
\midrule
\texttt{4u6v\_0P} & 0 & 28 & 0 \\
\texttt{4zfg\_0P} & 0 & 21 & 0 \\
\texttt{7uxl\_1P} & 0 & 25 & 0 \\
\texttt{8t58\_0P} & 1 & 26 & 1 \\
\bottomrule
\end{tabular}
\captionof{table}{The four islands short of a full top-5 (gate epitope RMSD
$\le \SI{2.5}{\angstrom}$); all are large. Full per-island counts in
\texttt{results/dual\_island\_gate\_summary.csv}.}\label{tab:shortislands}
\end{center}

\paragraph{Where accessibility is unavailable: crystal gaps.}
For \num{3840} designs --- one island each of three epitopes (\texttt{5eu7\_1P},
\texttt{5fhx\_1P}, \texttt{6ztr\_0P}, $\num{1280}$ designs apiece) --- the clash and cylinder
columns are null, with status \texttt{too\_few\_epitope\_pairs}. The cause is the same crystal
coverage constraint as in Section~\ref{sec:noab}: those islands' residues are unresolved in
the native complex, so there are no native C$\alpha$ anchors to superimpose the design onto,
and neither the real antibody clash nor the cylinder can be placed. These designs keep their
geometry and PAE (and so are still gated and ranked on those terms); they simply carry no
accessibility score.

\subsection{Selection}\label{sec:selection}
The composite scorer (Section~\ref{sec:composite}) reads this table directly. For the
per-island deliverable we select the top five per (\texttt{id}, \texttt{island\_index}) ---
per \emph{island}, not per epitope --- so each island is ranked on its own and the PepSeq
assay can deconvolve which island drives binding. Because this set has the real antibody
clash, whether to keep \texttt{af3\_n\_clash\_res}$\,=\,0$ as a hard pre-filter or fold the
cylinder term into the score is the one selection choice still open. [final selection and
top-5 designs to follow]
