\section{Open questions and next steps}\label{sec:open}

\begin{enumerate}
\item \textbf{Within-DP3 island stratification (the clean control).} We have split by island
count (Section~\ref{sec:islands}): single-island designs are distributionally easier on every
metric. What remains is to split by island \emph{size}, since the single-island epitopes are
also shorter on average, to separate the ``contiguous-12 is easier'' confound from a real
scaffolding-difficulty effect. The per-epitope island metadata is in \texttt{dp2.parquet}
(\texttt{epitope\_chunks}, \texttt{contig\_string}).

\item \textbf{Fit the soft-gate dials once the label has breadth.} The move off percentile is done:
the scorer is now the soft-gate (Section~\ref{sec:composite}), a sigmoid activation per metric, which
is absolute, saturating, and differentiable --- so it can be fit and a DP3-trained score transfers to
DP4. What remains is to \emph{choose the dials by data rather than by prior}. Most midpoints still sit at the
four-filter thresholds --- epitope PAE is the one already moved onto its data (midpoint \num{2.5}), a
first instance of the retuning this item calls for --- and the steepnesses $k$ and weights are set by
hand; the first experimental label
has arrived (the DP3 binding data, Section~\ref{sec:experimental}) and scaffolded designs do enrich over
the no-antibody baseline, so the premise holds. But that set cannot yet fix them: every assayed design
already passed the four filters (no failures to contrast), its pooled metric--binding correlation is a
between-antibody artifact, and only two of the eight antibodies are sampled deeply. The requirement for
DP4 is specific: assay designs spread across the metric space, \emph{including ones the current filters
reject}, so binding can be regressed on the metrics with real variance and across many epitopes. That is
what the metric-space sampling arm (C5) is built to provide. Accessibility already earns its heavy weight
twice over --- on DP3 the cylinder predicts the true antibody clash at AUC $0.93$
(Section~\ref{sec:noab}), and it is the single best predictor of \emph{experimental} binding, the only
metric whose correlation survives the between-antibody confound (Section~\ref{sec:whatpredicts}).

\item \textbf{Adopting the soft-gate scorer and retuning its epitope-PAE midpoint (done).} The shipped
library is selected under the soft-gate scorer --- its C1 members reproduce the
\texttt{antibody\_softgate} ranking exactly, so the move off the percentile scorer is realized in what
ships, not just in the code. One dial was then retuned: the epitope-PAE midpoint moved from 5 to
\num{2.5}, set from the intra-epitope PAE distribution rather than borrowed from the global threshold
(Section~\ref{sec:composite}). That shifted roughly one design in ten --- \num{107} of the \num{1120} C1
selections --- with counts and components unchanged, and the library was re-selected, re-assembled, and
re-encoded at that setting. The remaining open work is not re-selection but \emph{fitting} the dials to
data: the midpoints, steepnesses, and weights are still a prior, to be replaced by a fit against the DP4
binding readout (previous item).

\item \textbf{Confirm the canonical DP3 metrics builder.} Determine whether
\texttt{compute\_metrics\_mpnn.py} (currently archived) is the live RFD3+MPNN metrics
producer, and promote it into \texttt{episcaf\_analysis/} if so (see \texttt{docs/MIGRATION.md}).

\item \textbf{Validate the DP3 scoring path} end-to-end with \texttt{--preset antibody},
confirming \texttt{cylinder\_ca\_clashes} is the correct accessibility column there.

\item \textbf{Finish the 12-mer AlphaFold3 data migration} and verify landing counts
against the manifest.

\item \textbf{Pin down the crystal-gap dropouts.} We can already count them locally: 6M0J and
4WAT lose 15 and 40 tiles to unresolved crystal regions (Table~\ref{tab:tilecount}). What is
left is to confirm these against the cluster pre-filter status (the explicit
\texttt{crystal\_epi\_incomplete} entries) and to check whether the lost tiles fall in
contiguous runs, i.e.\ whether whole regions go un-scorable from gaps.
\end{enumerate}
