\section{The production run: per-island scaffolding of dual-island epitopes}\label{sec:perisland}
The island analysis (Section~\ref{sec:islands}) turns on a question only an experiment can
settle: scaffold each island of a dual-island epitope on its own and assay whether it still
binds. This run builds those reagents. It is the first component (``individual island'') of the
DP4 library outlined in \texttt{docs/DP4designIdeas.pdf} --- \num{1600} peptides at the shipped
top-\num{20} depth --- and one of several arms of DP4 aimed at its
Question~1, what makes a minimal epitope.

We define an \emph{island} as one fixed-coordinate \texttt{A}-segment of the RFdiffusion3
contig string. For example the dp2 contig \texttt{15-15/A1-16/57-57/A81-81/15-15} has two
islands, \texttt{A1-16} (16 residues) and \texttt{A81-81} (1 residue), separated by generated
scaffold of the given lengths. A \emph{dual-island epitope} is a ledger id with
\texttt{epitope\_chunks}~$=2$, i.e.\ exactly two such segments; the DP3 set contains
\num{46} of them. We take \emph{island size} to be the \texttt{A}-segment span in residues
(not the count of antibody-contact residues within it).

By design, each dual-island epitope contributes its two islands as separate
design targets, and we keep the top \num{20} designs per island under the composite scorer.
The island lengths vary widely --- from a single residue to \num{51} --- and Table~\ref{tab:dualisland}
lists both islands of all \num{46} epitopes so the run is auditable epitope by epitope. One
case needs a rule: an island of a single residue cannot be presented on its own. Rather than
drop the whole epitope, we simply \emph{skip the size-\num{1} island and still scaffold its
partner}. Five islands are skipped on this ground, one each in \texttt{2h32\_0P},
\texttt{3q1s\_0P}, \texttt{6o9i\_0P}, \texttt{7a3t\_0P}, and \texttt{7tzh\_0P} (marked in
Table~\ref{tab:dualisland}); every epitope still contributes at least one island. Of the
\num{92} islands across the \num{46} epitopes, \num{87} are scaffolded, giving \num{87}
individual-island targets; at the shipped top-\num{20} per island (after the 56-mAb exclusion) that
is \num{1600} shipped designs (component C2). The per-island
target list (both spans, sizes, and the scaffold/skip flag) and Table~\ref{tab:dualisland}
itself are generated by \texttt{episcaf\_analysis/dual\_island\_targets.py} into
\texttt{results/dual\_island\_targets.csv} and \texttt{manuscript/tables/}, the single
reproducible source for these counts.

Each island is scaffolded into a fixed \num{103}-residue construct --- the maximum peptide
length the PepSeq assay synthesizes. We call the stretches of generated scaffold lying
N-terminal and C-terminal to a fixed island its \emph{flanks}; a single-island contig is
therefore $\langle\text{N-flank}\rangle/\,\texttt{A}a\text{-}b\,/\langle\text{C-flank}\rangle$,
and the scaffold budget the flanks share is $103-\ell$ residues for an island of length
$\ell$. We note one unexplained convention difference: Lawson's original contigs were
\num{104} residues, not \num{103}, and the change to \num{103} has no recorded origin. This
single-island run adopts \num{103} directly; the whole-epitope C1 set is likewise built natively at
\num{103}, rather than trimmed from Lawson's \num{104}-mers (Section~\ref{sec:c1set}), so both production
runs are at the PepSeq length.

\paragraph{Contig sampling and run size.}
A \emph{contig} pairs the fixed island(s) with one concrete scaffold-length layout, and each
contig is sampled at \num{8} RFdiffusion3 backbones $\times$ \num{8} ProteinMPNN sequences
$=\num{64}$ designs (the per-contig depth throughout this project). The run size is thus set
by how many contigs we generate. We record how Lawson reached his \num{151232}-design DP3 set,
so the scale here is never mysterious: for each epitope he held the two islands fixed and swept
the scaffold layout, keeping the two flanks balanced ($|\text{N}-\text{C}|\le 1$) and stepping
the flank length over every integer from \num{1} to about \num{20}, which makes the
\emph{inter-island gap} sweep from $\sim\!\num{46}$ down to a floor near \num{8} (the islands
are never allowed to touch). This is a deterministic enumeration, not random sampling: it
yields a median of $\sim\!\num{41}$ contigs per epitope, so $\num{59}\,\text{epitopes}\times
41 \approx \num{2363}$ contigs and $\num{2363}\times\num{64}=\num{151232}$ designs. The
quantity his contigs actually vary is the spacing between the two islands.

Our single-island targets have no inter-island gap, so that lever does not exist; the only
scaffold freedom is the \emph{flank asymmetry} --- where along the \num{103}-mer the island
sits, i.e.\ the split of the $103-\ell$ budget into N- and C-flank. We therefore generate $V$
contigs per island by \emph{randomly} sampling $V$ distinct N-flank lengths over
$[1,\,103-\ell-1]$ (under a fixed seed, so the ledger is reproducible). This is an explicitly
different diversity axis from Lawson's --- scaffold position and topology, not inter-island
geometry --- which we name here rather than leave implicit. With $V$ variants the run is
\begin{equation}\label{eq:runsize}
\underbrace{\num{87}}_{\text{islands}}\times\,V\,\times\,\underbrace{\num{8}}_{\text{RFD3}}\times\,\underbrace{\num{8}}_{\text{MPNN}}
\;=\;\num{5568}\,V \ \text{designs}.
\end{equation}
We use $V=\num{20}$, giving $\num{1740}$ contigs and $\num{111360}$ designs --- the order of
the previous DP3 set ($V\approx\num{27}$ would reach its full $\sim\!\num{150000}$). $V$ is the
\texttt{--variants} dial of \texttt{episcaf\_pipeline/build\_dual\_island\_designs.py}. Each
RFdiffusion3 task emits \num{8} backbones (\texttt{n\_batches=1}), so the \num{8}-per-contig
depth is one task, not eight: the RFD3 step is \num{1740} array tasks (one per contig), each
yielding \num{8} backbones that ProteinMPNN then re-sequences \num{8} ways.

\input{tables/dual_island_targets}

\subsection{One metrics table for every decision}\label{sec:metrics}
With the run designed above and pushed through RFdiffusion3 $\rightarrow$ ProteinMPNN
$\rightarrow$ AlphaFold3, we extract every per-design metric \emph{once} into a single table
and make all downstream choices (composite scoring, top-\num{20} per island) from it, so the analysis
is never recomputed. This is our analog of Lawson's \texttt{dp2.parquet}: one row per design,
written by \texttt{scripts/stage05\_extract\_metrics.py} from the \texttt{04\_af3/outputs}
predictions, the per-island ledger (\texttt{results/dual\_island\_designs.csv}), and the native
AbDb complex PDBs.

\paragraph{Why a dedicated extractor.}
The repository already held two metric extractors, and neither fit this run directly.
\texttt{compute\_metrics.py} is the Lawson-validated antibody extractor, but its run mode is
keyed on \texttt{dp2}'s hashed tokens and the no-MPNN (RFD3$\rightarrow$AF3) layout.
\texttt{build\_12mer\_metrics.py} emits exactly the columns the composite scorer reads
(\texttt{epitope\_pae}, the cylinder terms) but parses the 12-mer naming and assumes no
antibody. Stage~05 reuses the \emph{validated primitives} of the first (backbone RMSD,
PAE, AlphaFold3 file discovery, the clash geometry) and mirrors the PAE decomposition and
cylinder calls of the second, but drives them from our per-island ledger.

The per-island set sits in a useful middle: these are the 46 dual-island \emph{mAb} epitopes,
so a solved antibody exists and the real DP3 clash filter \emph{is} available. We therefore
compute both accessibility signals --- the true antibody clash
(\texttt{af3\_n\_clash\_res}, against native chains B/C) \emph{and} the cylinder surrogate ---
which makes this run a second check of the surrogate against ground truth, on top of the DP3
check in Section~\ref{sec:noab}.

\paragraph{What each design row carries.}
The table groups into identity, geometry, confidence, and accessibility
(Table~\ref{tab:metricschema}). Geometry and confidence follow the four-filter definitions of
Section~\ref{sec:fourfilter}; the accessibility block carries the real clash and the cylinder
side by side.

\begin{center}
\begin{longtable}{lll}
\caption{Schema of the per-design metrics table (\texttt{metrics\_dual\_island.parquet}).
One row per AlphaFold3 design.}\label{tab:metricschema}\\
\toprule
column & meaning & role \\
\midrule
\endfirsthead
\toprule column & meaning & role \\ \midrule \endhead
\multicolumn{3}{l}{\emph{Identity (which design this is)}}\\
\texttt{id}            & epitope id (e.g.\ \texttt{2h32\_0P}) & select / scope \\
\texttt{island\_index} & which of the two islands (0/1)        & select group \\
\texttt{island\_size}  & island span (residues)                & strata \\
\texttt{contig\_id}    & contig within the epitope             & provenance \\
\texttt{variant}, \texttt{n\_flank}, \texttt{c\_flank} & flank-split placement & provenance \\
\texttt{seed}, \texttt{rep}, \texttt{dldesign} & RFD3 / MPNN replica indices & provenance \\
\texttt{predID}, \texttt{af3\_dir} & design stem, output path     & join key \\
\midrule
\multicolumn{3}{l}{\emph{Geometry (backbone RMSD, Kabsch-superposed)}}\\
\texttt{epitope\_chunk\_rmsd} & design vs AF3, epitope only ($\le 1$) & soft gate, 0.25 \\
\texttt{overall\_rmsd}        & design vs AF3, full scaffold ($\le 2$) & soft gate, 0.20 \\
\midrule
\multicolumn{3}{l}{\emph{Confidence (AlphaFold3)}}\\
\texttt{epitope\_pae}  & mean PAE over epitope residues   & 0.10 \\
\texttt{scaffold\_pae} & mean PAE over the scaffold        & diagnostic \\
\texttt{mean\_pae}     & global mean PAE ($<5$, four-filter) & diagnostic \\
\texttt{ptm}, \texttt{ranking\_score} & AF3 summary scalars & diagnostic \\
\midrule
\multicolumn{3}{l}{\emph{Accessibility (this run has both)}}\\
\texttt{af3\_n\_clash\_res} & real antibody clash count ($=0$) & \textbf{0.45} (accessibility) \\
\texttt{cylinder\_ca\_clashes}   & scaffold C$\alpha$ in the cylinder & surrogate (unweighted here) \\
\texttt{cylinder\_native\_aware} & same, native volume carved out     & surrogate \\
\texttt{native\_in\_cylinder}    & native non-epitope C$\alpha$ in cylinder & test \\
\bottomrule
\end{longtable}
\end{center}

\paragraph{Definitional choices (held to the DP3-validated convention).}
Three choices fix the numbers, and we keep each at the value that makes this set comparable to
DP3. (i) Both RMSDs are computed on the \emph{backbone} (N, C$\alpha$, C, O) and
Kabsch-superposed, as in \texttt{compute\_metrics.py}, not the C$\alpha$-only variant of the
12-mer extractor. (ii) The cylinder uses \texttt{exclude\_dist}\,$=\,\SI{1}{\angstrom}$, the
value adopted for the validated 12-mer preset (Section~\ref{sec:noab}). (iii) The epitope is
the \emph{whole island span} $[\,\texttt{n\_flank},\ \texttt{n\_flank}+\texttt{island\_size}\,)$
--- the residues stage~03 actually held \textsc{fixed} during ProteinMPNN --- not the
contacts-only subset stored in the ledger. The native-frame and design-frame epitope residues
pair one-to-one because each island A-segment is contiguous in both, which we verified holds
for all \num{1740} ledger contigs ($b-a+1=\texttt{island\_size}$).

\paragraph{How the metrics table is produced.}
The extractor caches each native complex once per epitope (not once per design) and shards the
$\sim$46 epitopes across worker processes:
\begin{verbatim}
python3 scripts/stage05_extract_metrics.py \
    --af3_out_dir  runs/dual_island_rfd3/04_af3/outputs \
    --mpnn_pdb_dir runs/dual_island_rfd3/03_mpnn/mpnn_pdbs \
    --ledger       results/dual_island_designs.csv \
    --native_dir   <AbDb cleaned complex dir> \
    --out          runs/dual_island_rfd3/05_analysis/metrics_dual_island.parquet \
    --workers 16
\end{verbatim}
It writes the parquet and a CSV alongside. A companion report,
\texttt{scripts/stage05\_summarize.py}, prints the breakdowns below and writes the per-island
summary \texttt{results/dual\_island\_gate\_summary.csv} (so these numbers are regenerable,
not pasted).

\subsection{The run's yield}\label{sec:metricsresults}
The full run scored \num{111322} designs across all 46 epitopes and 87 islands, every one with
a status of \texttt{ok}. The designs are good where it matters: median epitope-chunk RMSD is
\SI{1.20}{\angstrom} (best \SI{0.03}{\angstrom}) and median epitope PAE is $3.1$, while the
scaffold is predicted less confidently (scaffold PAE median $7.9$) --- the same
epitope-sharp / scaffold-loose split the decomposed metrics were built to exploit
(Section~\ref{sec:filtering}).

\paragraph{The design funnel.}
Table~\ref{tab:funnel} walks a design through each filter. Two thirds clear the composite
scorer's gate, and nearly half hold the epitope to within \SI{1}{\angstrom}. The strict DP3
four-filter, however, passes only $0.37\%$ --- driven down by the global PAE ($<5$) and the
zero-clash requirements, exactly the regime seen on DP3 ($<1\%$, Section~\ref{sec:islands}).
This is why per-island \emph{selection ranks by the composite score within the gate}, rather
than taking only four-filter passes, which would leave most islands with nothing to advance.

\begin{center}
\begin{tabular}{lrr}
\toprule
filter (status\,=\,ok, $n=\num{111322}$) & designs passing & \% \\
\midrule
composite scorer gate (epitope RMSD $\le \SI{2.5}{\angstrom}$) & \num{74061} & 66.5 \\
\quad epitope-chunk RMSD $\le \SI{1}{\angstrom}$              & \num{51653} & 46.4 \\
\quad overall RMSD $\le \SI{2}{\angstrom}$                    & \num{45531} & 40.9 \\
\quad global mean PAE $< 5$                                   & \num{25849} & 23.2 \\
\quad antibody clash $= 0$ (of \num{107482} with clash)      & \num{2510}  & 2.3 \\
\midrule
DP3 four-filter, all four (real clash) & \num{407} & 0.37 \\
\bottomrule
\end{tabular}
\captionof{table}{Per-design pass counts. Indented rows are the individual four-filter
thresholds (not nested); the last row requires all four at once. Regenerated by
\texttt{scripts/stage05\_summarize.py}.}\label{tab:funnel}
\end{center}

\paragraph{These are single-island rates, and the comparison to Lawson's DP3 is two-edged.}
It is tempting to read Table~\ref{tab:funnel} against the four-filter rates of Lawson's DP3 set,
but the two differ in \emph{two} ways at once: the backbone generator (RFD1 vs RFD3) \emph{and}
the epitope (Lawson's set is dominated by two-island epitopes; this run scaffolds single
islands). Table~\ref{tab:disentangle} holds each fixed in turn: columns 1--2 are the full 59 DP3
epitopes, column 3 the \num{46} dual-island epitopes (\num{87} islands) this run splits.
On the \emph{same} DP3 contigs, swapping RFD1 for RFD3 barely moves the rates --- and slightly
lowers most --- so the generator is not the story (the full distributions are in
Section~\ref{sec:rfd}). Scaffolding a \emph{single} island is: the epitope-RMSD pass rate
roughly triples ($14\% \to 46\%$) and the global-PAE rate climbs from $7\%$ to $23\%$, because
one small, contiguous island is far easier to hold rigidly and predict confidently than two
spatially separated patches (Section~\ref{sec:islands}). The one filter that moves the other way
is accessibility: the clash-free rate \emph{falls} ($8\% \to 4\% \to 2\%$ over all designs),
because a small epitope on a fixed \num{103}-mer leaves more non-epitope scaffold mass to intrude
on the antibody approach --- the scaffold-mass effect of Section~\ref{sec:noab}. That cap on
clash is why the all-four rate ($0.37\%$) lands \emph{below} Lawson's ($0.56\%$ over all designs,
$0.68\%$ over its valid ones) even though three of the four filters improve markedly.

\begin{center}
\begin{tabular}{lrrr}
\toprule
 & RFD1 / Lawson$^{*}$ & RFD3 / same & RFD3 / single \\
filter & DP3 contigs & DP3 epitopes & island (this run) \\
\midrule
epitope RMSD $\le \SI{1}{\angstrom}$ & 13.7 (16.6) & 14.0 & 46.4 \\
overall RMSD $\le \SI{2}{\angstrom}$ & 22.2 (26.9) & 16.4 & 40.9 \\
clash $=0$                           &  8.3 (10.1) &  3.9 &  2.3 \\
mean PAE $<5$                         &  1.9 (2.3)  &  7.1 & 23.2 \\
all four                             & 0.56 (0.68) & 0.50 & 0.37 \\
\bottomrule
\end{tabular}
\captionof{table}{Per-filter pass rates (\%) disentangling generator from island count. Columns
1--2 are the full \num{59}-epitope DP3 set; column 3 is the \num{46} dual-island epitopes split into
\num{87} single-island targets. Column 1$\to$2 isolates RFD1 vs RFD3 (identical contigs); column
2$\to$3 isolates two-island vs single-island (both RFD3). Each cell is the rate over \emph{all}
designs generated. $^{*}$The RFD1 parenthetical is the rate over only its \emph{valid} designs
(all four metrics present); the two diverge because Lawson's set is missing an AlphaFold3 result
for \num{26870} of its \num{151232} designs (its valid column reproduces the reference hand calculation,
17/27/10/2/0.7). The RFD3 columns show a single rate because those sets are essentially complete
(RFD3/same missing \num{12} of \num{150948}; RFD3/single missing only a clash for the \num{3840}
crystal-gap designs of \num{111322}), so their all- and valid-rates coincide. Regenerated by
\texttt{scripts/compare\_passrates.py}.}\label{tab:disentangle}
\end{center}

\paragraph{Most islands fill their depth; a few large ones produce almost nothing.}
The shipped depth is top-\num{20} per island. Even against a much shallower floor of five,
\textbf{83 of the 87 islands clear it}; the four that cannot (Table~\ref{tab:shortislands}) are
\emph{all large islands, 21--28 residues} --- the island-size difficulty of Section~\ref{sec:islands}
surfacing directly in the deliverable: the biggest islands are the hardest to scaffold cleanly, and
three of them produce no gated design at all. For these we advance whatever clears the gate (a handful,
or none) and say so, rather than padding the set with designs that miss the bar.

\begin{center}
\begin{tabular}{llrr}
\toprule
epitope & island & island size (aa) & gated designs (of \num{1280}) \\
\midrule
\texttt{4u6v\_0P} & 0 & 28 & 0 \\
\texttt{4zfg\_0P} & 0 & 21 & 0 \\
\texttt{7uxl\_1P} & 0 & 25 & 0 \\
\texttt{8t58\_0P} & 1 & 26 & 1 \\
\bottomrule
\end{tabular}
\captionof{table}{The four islands that clear the gate (epitope RMSD $\le \SI{2.5}{\angstrom}$) with fewer
than five designs --- far short of the shipped top-\num{20}; all are large. Full per-island counts in
\texttt{results/dual\_island\_gate\_summary.csv}.}\label{tab:shortislands}
\end{center}

\paragraph{Where accessibility is unavailable: crystal gaps.}
For \num{3840} designs --- one island each of three epitopes (\texttt{5eu7\_1P},
\texttt{5fhx\_1P}, \texttt{6ztr\_0P}, $\num{1280}$ designs apiece) --- the clash and cylinder
columns are null, with status \texttt{too\_few\_epitope\_pairs}. The cause is the same crystal
coverage constraint as in Section~\ref{sec:noab}: those islands' residues are unresolved in
the native complex, so there are no native C$\alpha$ anchors to superimpose the design onto,
and neither the real antibody clash nor the cylinder can be placed. These designs keep their
geometry and PAE (and so are still gated and ranked on those terms); they simply carry no
accessibility score.

\subsection{Selecting the per-island deliverable}\label{sec:selection}
The composite scorer (Section~\ref{sec:composite}) reads this table directly. For the
per-island deliverable we select the top \num{20} per (\texttt{id}, \texttt{island\_index}) ---
per \emph{island}, not per epitope --- so each island is ranked on its own and the PepSeq
assay can deconvolve which island drives binding. The policy is to \textbf{rank, not gate}: we
apply no hard threshold and take the top \num{20} designs per island by composite score, since the
composite already penalizes weak metrics softly and a hard cut would only discard designs the
ranking buries anyway. Because this is a known-antibody set, the accessibility term in the
composite is the \emph{real} clash \texttt{af3\_n\_clash\_res} rather than the cylinder surrogate
(Section~\ref{sec:composite}): a design that places mass where the antibody binds is pushed down
the ranking, but is not categorically excluded. This yields \num{1600} shipped C2 constructs of the
shipped library (\texttt{results/dp4\_C2\_single\_island\_ranked.top20.csv}).
